Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex

\section{AE Comments}

\pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}

\textcolor{teal}{Suggestion: Write a strong paragraph to the Editor after the paper is finalized. Come back to this. Explain that we have tried to simplify the notation when possible, however, in some places, the heavy notation is not possible. Thus, we have worked on improving more precise wording and removing sections or wording that is distracting or not needed. Thank the Editor and Associate Editor for helping us improve upon what we believe to be an important paper and breakthrough in the literature.}

\textcolor{teal}{I would not make a large list of all the notation we have changed as it may cause the Editor to go picking over the paper more.}

\reply{
We thank the Editor for the opportunity to revise our paper and for the suggestions. We have improved notation, when possible, however, as noted by the Editor in certain places, such as in the computational speeds ups of Section 4, this is difficult to avoid. We have done our utmost to simplify notation (when possible) and improve the parsing of sentences. Finally, we have attempted to make it more clear the general contributions of our paper, which we highlight for the Editor below.

We propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. {\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}. We employ the technique of \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.}
This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
}

%\reply{We have made several efforts to simplify notation in this revision.
%
%\begin{itemize}
%	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. This particularly simplifies the notation in Section 4.2.
%	
%	\item At the end of Section 4.1, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics.
%	This likelihood is then used throughout Section 4.2 in the discussion of efficient posterior inference. We hope this expression clarifies the meaning of "preservation of weight (Comment 5), and strengthens the argument for $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ being summary statistics for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ (Comment 4). 
%	
%	\item To denote when the $(i,j)$ record pair exhibits agreement pattern $p$, we have removed the $(i, j) \in h_p$ notation for $\gamma_{ij} = h_p$.
%	
%	\item In Section 4.1, we have provided a specific example of a comparison vector and how the hashing procedure works. 
%	
%	\item Section 4.1 introduces the summary statistic $N_{j_p}$. After careful review, we realized that by changing the order of subscripts and instead using $N_{p_j}$, we can replace the sum of counts across all records (previously denoted $H_p$), with the more intuitive $N_p$. Thus, we have removed all of the ''$H$" quantities throughout Section 4. 
%\end{itemize}
%}

@@

\reply{%
	$\gamma$ is a matrix comprised by comparison vectors. All references to $\gamma$ as a set have been removed.
}

\pointRaised{R}{% 
	Section 4.3, second paragraph, the definitions of the concatenated vectors $\alpha_0$ and $\beta_0$ should have final elements subscripted by $L_f$, where the sub-subscript is capitalized.}

\reply{%
	This has been revised, thank you
}

	\clearpage

	\bigskip

	\bibliographystyle{jasa}
	\bibliography{biblio}

### 2023-07-17 ba_resubmit/kundinger_resubmit2.tex
working on response to reviewers.

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{data fusion}
\kwd{distributed computing}
\kwd{entity resolution}
\kwd{hashing}
\kwd{record linkage}
\end{keyword}

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity. However, In many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

@@

In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. {\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

>>>>

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{bipartite record linkage}
\kwd{entity resolution}
\kwd{data cleaning}
\kwd{parallelization}
\kwd{hashing}
\end{keyword}

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity. However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}. For example, the decision theoretic work of \cite{steorts_bayesian_2016}, ensures that the resulting records are bipartite (or rather transitive).}

%but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

@@

In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. {\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}. We employ the technique of \cite{steorts_bayesian_2016} in order to ensure our method after record linkage is bipartite.}
This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.
